---
title: "Using Multiple Linear Regression To Predict La-Liga Standings"
author: "Debaditya Chakraborty"
date: "2023-09-12"
output: rmdformats::downcute
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

# Data Cleaning

We Start with loading the data from the excel file , the file can be accessed from [here](https://github.com/niceGriffith/La-liga-Prediction-Using-MLR/blob/main/Project_updated.xlsx).  
For this we use the packages `xlsx`(For .xlsx file handling),`dplyr` and`tidyverse` for data manipulation.
```{r}
library(xlsx)
library(dplyr)
library(tidyverse)
mk<-read.xlsx("Project_updated.xlsx",sheetIndex = 1)
#Clearing subcategory related empty rows,
mk<-mk[-1,]
# Assigning correct column names
colnames(mk)[4:6]<-c("Income","Expenditure","Gross")
#Removing some Columns which won't be used
mk<-mk[,-12:-10]
#Applying an algorithm to replace the team name rows 
## with NA team names with correct team names
mk <- mk %>%
  # Create a group identifier based on the row number
  mutate(group = (row_number() - 1) %/% 5) %>%
  # Group by the new group identifier
  group_by(group) %>%
  # Fill the missing values in the team column with the first non-missing value in each group
  fill(Team, .direction = "down") %>%
  # Ungroup and remove the temporary group column
  ungroup() %>%
  select(-group)
#Making data numeric
mk <- mk %>% mutate(across(-2:-1,as.numeric))
#Removing empty rows
mk<-mk[-153:-136,]
#Removing NA's
fin_mk<-na.omit(mk)
```

# Scatterplots

Here we plot some scatterplot between different variables in the dataset to get how they are related.

```{r}
par(mfrow=c(2,2))
plot(fin_mk$Average.Age.of.Players,fin_mk$Points.Scored,xlab = "Average Age of players",ylab="Points Scored")
plot(fin_mk$Expenditure,fin_mk$Points.Scored,xlab = "Expenditure",ylab="Points Scored")
plot(fin_mk$Manager.s.Years.of.experience,fin_mk$Points.Scored,xlab="Managers Year of Experience", ylab="Points Scored")
plot(fin_mk$Income,fin_mk$Points.Scored,xlab="Income", ylab="Points Scored")
plot(fin_mk$Average.Age.of.Players,fin_mk$Gross,xlab="Average Age of players",ylab="Financial Situation")
plot(fin_mk$Average.Age.of.Players,fin_mk$Manager.s.Years.of.experience,xlab="Average Age of players",ylab="Managers year of Experience")
plot(fin_mk$Average.Age.of.Players,fin_mk$Previous.Standing,xlab="Average Age of players",ylab="Previous Standings")
plot(fin_mk$Previous.Standing,fin_mk$Manager.s.Years.of.experience,xlab="Previous Standing",ylab="Managers year of Experience")
plot(fin_mk$Gross,fin_mk$Manager.s.Years.of.experience,xlab="Financial Situation",ylab="Managers year of Experience")
plot(fin_mk$Previous.Standing,fin_mk$Gross,xlab="Previous Standing",ylab="Financial Situation")
```

#  Histogram of Points Scored

```{r}
mk$Points.Scored%>%hist(main="Histogram of Points Scored")
```

# Correlation Matrix
We plot the correlation matrix using the `corrplot` packages corrplot function which is used to calculate the correlation matrix, and this is displayed via a visualization. 
```{r}
mk_num<-fin_mk[,-2:-1]
cor_mat<-cor(mk_num)
corrplot::corrplot(cor_mat)
```

# Modeling 

Finally we model the data using a multiple linear regression model, we use the explanatory variables Average age of players,Managers Year of Experience,Expenditure and Previous Standing.    
```{r}
model<-lm(Points.Scored~Average.Age.of.Players+Manager.s.Years.of.experience+Expenditure+Previous.Standing,fin_mk)
kk<-round(predict(model),digits = 0)
df1<-data.frame(fin_mk$Team,kk)
colnames(df1)<-c("Team","Points")
summary(model)
```

We can see that the R^2 value is particularly good,and we have to statistically significant explanatory variables in the model. Although this is a very overly simplified model but as a basis this looks decent.

# Prediction

Here the prediction is based on the average of the given data, i was too lazy to look for a 23-24 data ;__; 

If you want you can definitely collect the data and use this code to check how the model aligns with last years situation.
``` {r}
team_means <- df1 %>%
  group_by(Team) %>%
  summarize(mean_value = round(mean(Points, na.rm = TRUE),digits = 0))
standings<-team_means%>%arrange(desc(mean_value))
colnames(standings)[2]<-"Predicted Points"
library(knitr)
library(kableExtra)

kable(standings,caption = "Predicted Standings")

```

Thank you for reading.